Quantitative Auditing of Fairness Measures with Differentially Private Synthetic Data

### 2024-11-30 paper/paper.tex
removed ratios

Finally, we analyze the discrepancies between the fairness properties of the original and the synthetic by calculating the difference and the ratio of their perspective fairness measure values. The sum of the difference and the average of the ratio serve as a summary of the analysis.

@@

TODO: dataset sizes

We looked at several publicly available datasets, such as adult\cite{adult_2,Kaggle_Adult_Census_Income}, COMPAS\cite{larson2016propublica,Kaggle_COMPAS_Dataset}, and diabetes\cite{diabetes_34,Kaggle_Diabetes_Prediction}. The adult dataset comes from the 1994 census in the United States. The COMPAS dataset comes from an investigative report by ProPublica of the COMPAS criminal recidivism assessment system. The diabetes dataset comes from the hospital readmission data published in the 1994 AI in Medicine journal.

@@

We observed that across all examined fairness measures, their difference all fall below 0.1. The sum of their differences is 0.259 and the average of their ratios is 0.952, which we consider quite satisfactory.

\begin{table}[h]
\caption{Fairness measures experiment results of the adult dataset. Sum of differences is 0.259. Average of ratios is 0.952.}
\label{tab:adult_score}
\begin{tabular}{ccccc}
\toprule
\textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Diff} & \textbf{Ratio} \\
\midrule
Demographic Parity  & 0.172 & 0.104 & 0.067 & 1.651 \\
Accuracy Eqaulity   & 0.047 & 0.117 & 0.069 & 0.404 \\
Equalized Odds 1    & 0.057 & 0.079 & 0.021 & 0.723 \\
Equalized Odds 2    & 0.166 & 0.122 & 0.044 & 1.365 \\
Accuracy Equality 1 & 0.100 & 0.132 & 0.031 & 0.759 \\
Accuracy Equality 2 & 0.119 & 0.146 & 0.027 & 0.814 \\
\bottomrule
\end{tabular}
\end{table}

@@

The results showed an increase in error on the sufficiency measure values. In particular, one of the measures has an error ratio as high as 6. The sum of their differences is 0.419, and the average of their ratios is 1.745, which is worse than the adult dataset.

\begin{table}[h]
\caption{Fairness measures experiment results of the COMPAS dataset. Sum of differences is 0.419. Average of ratios is 1.745.}
\label{tab:compas_score}
\begin{tabular}{ccccc}
\toprule
\textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Diff} & \textbf{Ratio} \\
\midrule
Demographic Parity     & 0.131 & 0.098 & 0.032 & 1.329 \\
Accuracy Eqaulity      & 0.007 & 0.013 & 0.005 & 0.575 \\
Equalized Odds 1       & 0.024 & 0.099 & 0.074 & 0.249 \\
Equalized Odds 2       & 0.017 & 0.097 & 0.079 & 0.182 \\
Accuracy Equality 1    & 0.170 & 0.082 & 0.088 & 2.082 \\
Accuracy Equality 2    & 0.169 & 0.027 & 0.141 & 6.058 \\
\bottomrule
\end{tabular}
\end{table}

@@

There was also an increase of error on some sufficiency measure values. The sum of their differences is 0.189 and the average of their ratios is 0.489. However, this comparison may be skewed because one of the fairness measures in the original data was calculated to be zero, thus making the average ratio seem lower than otherwise.

\begin{table}[h]
\caption{Fairness measures experiment results of the COMPAS dataset. Sum of differences is 0.189. Average of ratios is 0.489.}
\label{tab:diabetes_score}
\begin{tabular}{ccccc}
\toprule
\textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Diff} & \textbf{Ratio} \\
\midrule
Demographic Parity  & 0.013 & 0.015 & 0.002 & 0.867 \\
Accuracy Eqaulity   & 0.007 & 0.009 & 0.002 & 0.782 \\
Equalized Odds 1    & 0.000 & 0.003 & 0.003 & 0.000 \\
Equalized Odds 2    & 0.008 & 0.106 & 0.097 & 0.081 \\
Accuracy Equality 1 & 0.013 & 0.097 & 0.084 & 0.136 \\
Accuracy Equality 2 & 0.021 & 0.020 & 0.001 & 1.068 \\
\bottomrule
\end{tabular}
\end{table}

TODO: only compare diff. remove ratios. use avg of diff.

>>>>

Finally, we analyze the discrepancies between the fairness properties of the original and the synthetic by calculating the difference of their perspective fairness measure values. The average of the differences serves as a summary of the analysis.

@@

We looked at several publicly available datasets, such as adult\cite{adult_2,Kaggle_Adult_Census_Income}, COMPAS\cite{larson2016propublica,Kaggle_COMPAS_Dataset}, and diabetes\cite{diabetes_34,Kaggle_Diabetes_Prediction}. The adult dataset comes from the 1994 census in the United States and contains about 30000 individuals. The COMPAS dataset comes from an investigative report by ProPublica of the COMPAS criminal recidivism assessment system and contains about 7000 individuals. The diabetes dataset comes from the hospital readmission data published in the 1994 AI in Medicine journal and contains about 100000 individuals.

@@

We observed that across all examined fairness measures, their difference all fall below 0.1. The average of their differences is 0.0431, which we consider quite satisfactory.

\begin{table}[h]
\caption{Fairness measures experiment results of the adult dataset. Average of differences is 0.0431.}
\label{tab:adult_score}
\begin{tabular}{ccccc}
\toprule
\textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Difference} \\
\midrule
Demographic Parity  & 0.172 & 0.104 & 0.067 \\
Accuracy Eqaulity   & 0.047 & 0.117 & 0.069 \\
Equalized Odds 1    & 0.057 & 0.079 & 0.021 \\
Equalized Odds 2    & 0.166 & 0.122 & 0.044 \\
Accuracy Equality 1 & 0.100 & 0.132 & 0.031 \\
Accuracy Equality 2 & 0.119 & 0.146 & 0.027 \\
\bottomrule
\end{tabular}
\end{table}

@@

The results showed an increase in error on the sufficiency measure values. In particular, one of the measures has an error difference as high as 0.141. The average of their differences is 0.069, which is worse than the adult dataset.

\begin{table}[h]
\caption{Fairness measures experiment results of the COMPAS dataset. Average of differences is 0.069.}
\label{tab:compas_score}
\begin{tabular}{cccc}
\toprule
\textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Difference} \\
\midrule
Demographic Parity     & 0.131 & 0.098 & 0.032 \\
Accuracy Eqaulity      & 0.007 & 0.013 & 0.005 \\
Equalized Odds 1       & 0.024 & 0.099 & 0.074 \\
Equalized Odds 2       & 0.017 & 0.097 & 0.079 \\
Accuracy Equality 1    & 0.170 & 0.082 & 0.088 \\
Accuracy Equality 2    & 0.169 & 0.027 & 0.141 \\
\bottomrule
\end{tabular}
\end{table}

@@

There was also an increase of error on some sufficiency measure values. The average of their differences is 0.031.

\begin{table}[h]
\caption{Fairness measures experiment results of the COMPAS dataset. Average of differences is 0.031.}
\label{tab:diabetes_score}
\begin{tabular}{cccc}
\toprule
\textbf{Measure} & \textbf{Original} & \textbf{Synthetic} & \textbf{Difference} \\
\midrule
Demographic Parity  & 0.013 & 0.015 & 0.002 \\
Accuracy Eqaulity   & 0.007 & 0.009 & 0.002 \\
Equalized Odds 1    & 0.000 & 0.003 & 0.003 \\
Equalized Odds 2    & 0.008 & 0.106 & 0.097 \\
Accuracy Equality 1 & 0.013 & 0.097 & 0.084 \\
Accuracy Equality 2 & 0.021 & 0.020 & 0.001 \\
\bottomrule
\end{tabular}
\end{table}
